---
title: "AbemaImmune"
author: "Corinne Strawser"
date: "November 14, 2024"
title-block-banner: true
format:
    html:
        toc: true
        toc_float: true
        toc_depth: 5
        number_sections: false
        code-fold: true
        code-link: true
        df-print: kable
        self-contained: true
execute:
    cache: true
knitr:
    opts_chunk:
        cache.lazy: false
        echo: FALSE
        error: FALSE
        cache: TRUE
        message: FALSE
        warning: FALSE
lightbox: true
---

# Sample summary

# Technical details

# Libraries
```{r}
library(Seurat)
library(tidyverse)
library(clustree)
library(RColorBrewer)
library(msigdbr)
library(fgsea)
library(data.table)
library(pheatmap)
library(viridis)
library(patchwork)
library(clusterProfiler)
library(enrichplot)
library(matrixStats)
library(survival)
library(survminer)
library(scales)
```

# Directories
```{r Dirs}
data_dir <- "~/projects/240320_abemaimmune/output/objects/"
fig_dir <- "~/projects/240320_abemaimmune/manuscript/figures/"
tbl_dir <- "~/projects/240320_abemaimmune/manuscript/tables/"
dir <- "~/projects/240320_abemaimmune/manuscript/"
```

# Colors
```{r Colors}
colors <- c("#0166d8", "#ec102f", "#3aa609", "#6524ff", "#f8d147", "#5ce8ef", "#fd6ca0",
            "#476ba5", "#fea27a", "#add465", "#b775e1", "#fe8f06", "#5cfdc4", "#feafda",
            "#00228b", "#fd5925", "#166d2a", "#a90aa1", "#f3c011", "#5ac7dc", "#c1c2f5",
            "#1f29a1", "#e13224", "#86ec5a", "#672396", "#e1c637", "#069668", "#fcb2c7",
            "#1288da", "#c26a6b", "#6e9d50", "#1e0e76", "#fea53b", "#15729c", "#ffa8ff",
            "#270fe2", "#8a1341", "#285d28", "#956bed", "#f79302", "#116966", "#d945c2",
            "#94a6fd", "#e13219", "#01c472", "#afc6fe", "#e3c60b", "#aedfca", "#c2639f",
            "#0e2450", "#6d1a44", "#91fc40", "#5858e1", "#fca283", "#1eefc9", "#fb2076", 
            "#8bd0eb")
patients <- c("#FFC107", "#449AE4", "#78B389", "#B32357", "#e13219", "#14C7BA", "#871C9A",
              "#FD8305", "#3B3EDE", "#E884AF", "#285d28", "#afc6fe", "#aedfca", "#2C2E9C")
two <- c("#1288da", "#fd6ca0")
```

# Full data set analysis and subsetting (Figure 1 and Supplemental Figure 1)

{{< include full.qmd >}}

# Lymphoid data set analysis (Figure 2 and Supplemental Figure 2-3)

{{< include lymphoid.qmd >}}

# Myeloid data set analysis (Figure 3 and Supplemental Figure 5-7)

{{< include myeloid.qmd >}}

# Comparison of myeloid cell clusters 0 and 5 (Supplemental Figure 8)

{{< include myeloid_cluster0_vs_cluster5.qmd >}}

# Lymphoid and myeloid cell type signature analysis (Supplemental Figure 4 and Fig 4)

{{< include signature.qmd >}}

# Environment Info

```{r sessionInfo}
sessionInfo()
```

